---
title: "Lab 1"
author: "Larbi Moukhlis"
output:
  html_document:
    code_download: true
    toc: true
  pdf_document:
    toc: true
---

# Make sure to add remaining questions from the PDF posted to Sakai on the assignment. 

### Q1: Create variables for the following and assign the value to the right of the variable name for each:
- Firstname Zach
- Lastname Lavine
- Position SG
- Age 25
- Team Chicago Bulls
- PPG 24.9
- AllStarYear 2024
    
```{r}
Firstname = "Zach"
Lastname = "Lavine"
Position = "SG"
Age = 25
Team = "Chicago Bulls"
PPG = 24.9 
AllStarYear = 2024
```


### Q2: Display the results and check the class of each of the variables (variables: Firstname, Lastname, Position, Age, Team, PPG, AllStarYear)
- display the results of each variable (what is the value)
- display the class of each (what is the data type) class()
```{r}
Firstname
Lastname
Position
Age
PPG
AllStarYear

class(Firstname)
class(Lastname)
class(Position)
class(Age)
class(PPG) 
class(AllStarYear)
```


### Q3: Change the data types as listed--if needed--based on class results of last chunk
- Firstname char (character)
- Lastname char (character)
- Position factor (factor)
- Age int (integer)
- Team factor (factor)
- PPG num (numeric)
- AllStarYear factor (factor)
- Rerun the class on each and display the results after any changes

```{r}
Position = as.factor(Position)
Age = as.integer(Age)
AllStarYear = as.factor(AllStarYear)
```

### Q4: Remove the variables created from the Global Environment
- rm(variablename)
- Firstname, Lastname, Position, Age, Team, PPG, AllStarYear
```{r}
rm(Firstname,Lastname, Position, Age, Team, PPG, AllStarYear)
```

### Q5: Create a vector for each element/column below; display the vectors, display the data type and length of each;
firstname lastname Position Age Team PPG AllStarYear
Zach Lavine SG 24 Bulls 24.9 2024
Kevin Durant F 33 Nets 29.3 2024
Demar DeRozan G-F 32 Bulls 26.4 2024
Trae Young G 23 Hawks 27.7 2024
- Required for Question 5:
- Create a vector for each column (name the vector the column name above)
- display the class for each vector class()
- display the length of each vector length()

```{r}
Firstname = c("Zach", "Kevin", "Demar", "Trae")
Lastname = c("Lavine", "Durant", "DeRozan", "Young")
Position = c("SG", "F", "G-F", "G")
Age = c(24,33,32,23)
Team = c("Bulls", "Nets", "Bulls", "Hawks")
PPG = c(24.9, 29.3,26.4,27.7)
AllStarYear = c(2024,2024,2024,2024)
```

### Q6: Create the dataframe called AllStarsEast; display the dataframe
- (firstname, lastname, Position, Age, Team, PPG, AllStarYear)
```{r}
AllStarEast = data.frame(Firstname, Lastname, Position, Age, Team, PPG, AllStarYear)

```

### Q7: Run str() to see the types of the dataframe columns str(AllStarsEast
```{r}
str(AllStarEast)

```

### Q8: Change any column needed to the following data types AND run str again on AllStarsEast. Make sure to use the
‘dataframe prefix’ on both sides before the columns names.
- firstname char (character)
- lastname char (character)
- Position factor (factor)
- Age int (integer)
- Team factor (factor)
- PPG num (numeric)
- AllStarYear factor (factor)

```{r}
Position = as.factor(Position)
Age = as.integer(Age)
Team = as.factor(Team)
AllStarYear = as.factor(AllStarYear)
str(AllStarEast)
```

### Q9: Display the average PPG by Team for dataframe AllStarsEast | Hint: tapply(AllStarsEast$PPG,
AllStarsEast$Team, mean) 
```{r}
tapply(AllStarEast$PPG, AllStarEast$Team, mean)
```

### Q10: Display the average PPG by Position for dataframe AllStarsEast | Hint: tapply tapply(AllStarsEast$PPG,
AllStarsEast$Position, mean) 
```{r}
tapply(AllStarEast$PPG,AllStarEast$Position,mean)
```

### Q11: Display the frequency for Team in dataframe AllStarsEast | Hint: table(AllStarsEast$Team)
```{r}
table(AllStarEast$Team)
```

### Q12: Answer the following questions descriptive statistic methods:
- Run code to find the average age. Comment the response to the following question using the code output: What is the average age from dataframe AllStarsEast?
- Run the summary statistics for the entire AllStarsEast dataframe.
**Notice the difference for character vs factor data types in the output**
```{r}
mean(AllStarEast$Age)
summary(AllStarEast)
```

### Q13: Use dataframe[Rows, Columns] to display the first row of data for all columns.
```{r}
AllStarEast[1,]
```

### Q14: Use dataframe[Rows, Columns] to display the points per game, PPG (column 6) of data for the 3rd player (row 3).
```{r}
AllStarEast[3,6]
```

### Q15: Use dataframe[Rows, Columns] to replace the points per game, PPG (column 6) of datafor the 3rd player (row 3) to the value of 27
```{r}
AllStarEast[3, 6] <-27
```

### Q16: Create a subset from AllStarsEast to only include the players from the team Bulls and call this Bullsplayers
```{r}
Bullsplayers <- subset(AllStarEast, Team == "Bulls")
```

### Q17: Display the new data frame, Bullsplayers
```{r}
print(Bullsplayers)
```



### SAVING and CREATING html FILE
After completing your R Notebook or R Markdown file, you will execute all the chunks, save the file, and generate the html file containing the codes and the results. 

