---
title: "Lab 6"
author: "Larbi Moukhlis"
date: "26 March 2025"
output: 
  html_document: 
    code_download: yes
    toc: yes
---


**Copyright Kaitlyn Hoffmann, All rights reserved. Not for resale, distribution, or posting to external sites for all content provided.**

# Classification: Decision Trees
#-----------------------------------------------

**IMPORTANT NOTE: do not put install.package code into a chunk or you may not be able to knit. To run, select the two lines and click Run-- Run selected line(s)**
#-----------------------------------------------

## Load/Call each library (rpart and rpart.plot). Needs to be called each time you need to access the functions rpart or rpart.plot
```{r}
library (rpart)
library (rpart.plot)
```

## Question 1: Create a data frame loaded from fetal_health.csv and call it fetal. 
**Note: there is a column called fetal_health that we will be using (do not get it confused later on)**
```{r}
fetal = read.csv("fetal_health.csv")
```

## Question 2: Display the first 10 rows from the data frame called fetal
```{r}
head(fetal, 10)
```



## No Code needed, just run: Create a subset from fetal with 70% of the rows assigned to a train dataframe and the remaining 30% assigned to a test data frame
```{r}
## 70% of the sample size
smp_size <- floor(0.70 * nrow(fetal))
## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(fetal)), size = smp_size)
train <- fetal[train_ind, ]
test <- fetal[-train_ind, ]
```


**Now, you will use train and test for the following: using fetal (onward) will result in loss of points**


## Question 3: Run the decision tree function. This will generate the algorithm (decision tree) to predict the fetal_health based on the characteristics of the patients (age, sex, blood pressure, cholesterol, Sodium).
```{r}
fetal_tree = rpart(fetal_health~., fetal, method="class")
```

## Question 4: Display the decision tree for predicting the fetal_health
```{r}
rpart.plot(fetal_tree, type=4, extra=101)

```


## Question 5: Display the frequency: fetal_health from the validation data
```{r}
table(test$fetal_health)
```

## Question 6: Predict the fetal_health for patients in the validation data and call the predict vector p
```{r}
vector_p = predict(fetal_tree, test, type="class")
```


## Question 7: Add the predicted_health as a new column to the test data frame (validation data) using the vector p from the previous chunk
```{r}
test$predicted_health = vector_p
```


## Question 8: Display the frequency: look at the actual (fetal_health) and the predicted fetal_health from validation
-   What column is the original fetal_health (what index)?
-   What column is the new predicted_health (what index)?
```{r}
# What column is the original fetal_health (what index)?: 22
# What column is the new predicted_health (what index)?: 23
#Now, use the two columns above in the rows,columns for the table() using the data frame test
table(test[,22:23])
(477+64+54)/nrow(test)
```


## Question 9: Reading the table results:    
```{r}
#Result: 595 out of 638 are correctly predicted
#The accuracy level =  595 / 638 = 93.27% 
```



## Question 10: Just as we did in class, walkthrough the following rows and comment what the predicted classification should be (look at the decision tree). In other words, each row represents a fetal patient, tell me what fetal_health would be administered to each. Comment your responses. There is an image of the decision tree in the homework pdf for this assignment on Sakai (check the assignment)
-   Row 1: mean_value_of_short_term_variability = 0.60 | histogram_mean = 110 | prolongued_decelerations = 0.001 | percentage_of_time_with_abnormal_long_term_variability = 8 | histogram_max = 190 | baseline_value = 135
-   Row 2: mean_value_of_short_term_variability = 0.52 | percentage_of_time_with_abnormal_long_term_variability = 65 | abnormal_short_term_variability = 85 | uterine_contractions = 0.005 | histogram_max = 160
-   Row 3: mean_value_of_short_term_variability = 0.58 | histogram_mean = 115 | prolongued_decelerations = 0.0005 | percentage_of_time_with_abnormal_long_term_variability = 6 | histogram_max = 180 | baseline_value = 125

The answer is either 1, 2, or 3.

```{r}
#Row1: 1
#Row2: 2 
#Row3: 3
```

### Knit to HTML





**Copyright Kaitlyn Hoffmann, All rights reserved. Not for resale, distribution, or posting to external sites for all content provided.**



