---
title: "Lab_3"
author: "Larbi Moukhlis"
output: 
  html_document: 
    code_download: yes
    toc: yes
---
**NOTE: do not change any code above other than the name, do not add the date to line 4--it auto populates the date)**

# R Markdown

----------------------------------------------------------------------------------

## Question 1: Load in the data starbucksmenu.csv as a data frame object called sbux
```{r}
sbux = read.csv("starbucksmenu.csv")
```

## Question 2: Display the first 8 rows of the data frame object sbux
```{r}
head(sbux, 8) 
```

## Question 3: Display the first 10 rows of the data frame object sbux
```{r}
head(sbux, 10)
```

## Question 4: Display the last 6 rows of the data frame object sbux
```{r}
tail(sbux, 6)
```


## Question 5: Display the structure of the data frame sbux; look at the data types and make sure they are loaded appropriately
-   display the structure (should only be 3 chr characters)
-   Bucket the categorical vs numerical variables/columns. Comment your response using # within the chunk determining whether each column/variable is numerical or categorical.
```{r}
#str (code)
str(sbux)

#categorical: Beverage_category, Beverage, Beverage_Prep
#numerical: Calories, Total_Fat_g, Trans_Fat_g, Saturated_Fat_g, Sodium_mg, Total_Carbs_g, Cholesterol_mg, Fiber_g, Sugar_g, Protein_g, Vitamin_PercDv, VitaminC,_PercDv, CalciumPercDv, Iron_PercDv, Caffeine_mg 
```

## Question 6: Display the total number of rows and columns in the data frame sbux
-   what does the number of rows represent (what is the dataset—what does each row represent..familiarize)? Comment your response using # within the chunk.
```{r}
#number of rows (code)
nrow(sbux)
#number of columns (code)
ncol(sbux)

# what does the number of rows represent (conceptually)?:
# Each row in the data frame represents a unique Starbucks drink with its own nutritional facts and preparation style
```

## Question 7: Basic calculations (answer using the code for each bullet, like mean/sd/max etc.):
- What is the average Calories from sbux?
- What is the average Caffeine_mg from sbux?
- What are the max Calories from sbux?
- What is the spread/standard deviation (variation) Calories from sbux?
- What is the range for Calories from sbux?
```{r}
mean(sbux$Calories)
mean(sbux$Caffeine_mg)
max(sbux$Calories)
sd(sbux$Calories)
range(sbux$Calories)
```

## Question 8: Display the summary results for the data frame called sbux
- What is the average calories for the drinks? Comment your response using # within the chunk
```{r}
summary(sbux)
# 198
```

## Question 9: Display the summary results for each (BY) Beverage_category separately for column 4 (Calories) only
(points off if not correct view)
- Analyze the results
- Which beverage category has the highest calories on average (mean)? Comment your response using # within the
chunk.
- Which beverage category has the lowest calories on average (mean)? Comment your response using # within the
chunk.
- Which beverage category has the highest calories (max)? Comment your response using # within the chunk.
```{r}
by(sbux[,4],sbux$Beverage_category,summary)
# Frappucino 
# Coffee
# Espresso
```

## Question 10: Display a frequency table for Beverage_category from the data frame sbux
- How many Classic Espresso Drinks are available on the menu? Comment your response using # within the chunk.
```{r}
table(sbux$Beverage_category)
# 98
```

## Question 11: Display a table summarizing the Average (mean) Caffeine_mg by Beverage_category from the data frame
sbux
- Which category has the highest overall average mg of caffeine? Comment your response using # within the chunk.
- Which category has the lowest overall average mg of caffeine? Comment your response using # within the chunk.
```{r}
tapply(sbux$Caffeine_mg,sbux$Beverage_category, mean)
# Coffee 
# Tea
```

## Question 12:
- Display the summarized standard deviation (sd) Caffeine_mg by Beverage_category for sbux
- Display the summarized average (mean) Calories by Beverage_category for sbux
```{r}
tapply(sbux$Caffeine_mg,sbux$Beverage_category, sd)
tapply(sbux$Calories, sbux$Beverage_category,mean)
```

## Question 13: Display/Plot the distribution for the Calories using the data frame sbux (histogram)
- add a color col="darkseagreen4"
- add a title main = "Calorie Distribution Table for Starbucks Drinks"
```{r}
hist(sbux$Calories, main = "Calorie Distribution Table for Starbucks Drinks", col = "darkseagreen4")
```

## Question 14: Display a boxplot for Calories from the data frame sbux
- add a color col = "limegreen"
- add a title main = "Calorie Distribution for Starbucks Drinks"
- Based on the output, how many outliers are in the data? Comment your response using # within the chunk.
- Approximately (just look and provide the approximate), what is the first quartile? What is the second quartile? Comment
your response using # within the chunk.
```{r}
boxplot(sbux$Calories, main = "Calorie Distribution for Starbucks Drinks", col = "limegreen")
# One outlier
# 1st quartile appears to be around 100 calories 
# 2nd quartile appears to be around 200 calories
```

## Question 15: Display a boxplot for Calories by Beverage_category from the data frame sbux
- add a color col = "yellowgreen"
- add a title main = "Calorie Distribution by Beverage Category for Starbucks Drinks"
```{r}
boxplot(sbux$Calories~sbux$Beverage_category, main = "Calorie Distribution by Beverage Category for Starbucks Drinks", col = "yellowgreen")
```

## Question 16: Display a boxplot for Caffeine_mg by Beverage_category from the data frame sbux
- add a color col = "lightgreen"
- add a title main = "Caffeine(mg) Distribution by Beverage Category for Starbucks Drinks"
- Does Coffee or Espresso have higher caffeine? Comment your response using # within the chunk.
```{r}
boxplot(sbux$Caffeine_mg~sbux$Beverage_category, main = "Caffeine(mg) Distribution by Beverage Category for Starbucks Drinks", col = "lightgreen")
# Coffee
```

## Question 17: Display a scatterplot for Calories (x) and Sugar_g (y) from sbux [Hint: plot(x, y)]
- add a color col = "blue"
- add a title main = "Starbucks: Calories and Sugar"
- Does Calories and Sugar have a positive relationship/trend? Comment your response using # within the chunk. Yes or No
```{r}
plot(sbux$Calories, sbux$Sugar_g, main = "Starbucks: Calories and Sugar", col = "blue")
# Yes
```

## Question 18: Use par(mfrow=c(2,2)) to display the following plots in 2 row 2 columns on the output. The first line of code
in the chunk will be the par(mfrow=c(2,2)) and then within the same chunk, put the 4 lines of code for the plots below
directly on the next line of the par() code above. In this chunk, you will have 5 lines of code.
- Display a scatterplot for Calories (x) and Sugar_g (y) from sbux (col = "blue")
- Display a scatterplot for Calories (x) and Total_Carbs_g (y) from sbux (col = "palegreen3")
- Display a histogram distribution for Calories from sbux (col = "seagreen4")
- Display a boxplot for Caffeine_mg by Beverage_category from the data frame sbux (col = "limegreen")
```{r}
par(mfrow=c(2,2))
plot(sbux$Calories, sbux$Sugar_g, main = "Starbucks: Calories and Sugar", col = "blue")
plot(sbux$Calories, sbux$Total_Carbs_g, main = "Starbucks: Calories and Total Carbs", col = "palegreen3")
hist(sbux$Calories, main = "Calorie Distribution for Starbucks Drinks", col = "seagreen4")
boxplot(sbux$Caffeine_mg ~ sbux$Beverage_category, main = "Caffeine(mg) Distribution by Beverage Category", col = "limegreen")
```


**ADD THE QUESTIONS BELOW AND A CHUNK FOLLOWING EACH FROM THE PDF POSTED TO THE ASSIGNMENT**




### Knit to HTML

**Copyright Kaitlyn Hoffmann, All rights reserved. Not for resale, distribution, or posting to external sites for all content provided.**

