---
title: "Lab 2: Customer Sales Data Analysis"
author: "Your Name Here"
output: 
  html_document: 
    code_download: yes
    toc: yes
---

# Lab 2 

### Q1: Create variables for a single transaction
Create and assign values to the following variables:
-   CustomerID  1023
-   CustomerName  "Alice Johnson"
-   Product  "Laptop"
-   Quantity  2
-   UnitPrice  899.99
-   TotalPrice = Quantity * UnitPrice

```{r}
CustomerID = 1023
CustomerName = "Alice Johnson"
Product = "Laptop"
Quantity = 2 
UnitPrice = 899.99 
TotalPrice = Quantity * UnitPrice
```

### Q2: Display variables and check data types
-   to display, just type each object name. 
-   use class() to see the data type for each
```{r}
CustomerID
CustomerName
Product
Quantity 
UnitPrice
TotalPrice 
class(CustomerID)
class(CustomerName)
class(Product)
class(Quantity)
class(UnitPrice)
class(TotalPrice)


```

### Q3: Convert data types, if necessary ( I am saying what it should be in the instructions below. Check the previous chunk output to see what changes are necessary)
-   CustomerID  integer
-   CustomerName  character
-   Product  factor
-   Quantity  integer
-   UnitPrice  numeric
-   TotalPrice numeric
```{r}
CustomerID = as.integer(CustomerID)
Product = as.factor(Product)
Quantity = as.integer(Quantity)
```

### Q4: Create a data frame for multiple transactions: 
- Create a vector for each column:

| CustomerID | CustomerName    | Product   | Quantity | UnitPrice | 
|------------|---------------|------------|----------|----------|
| 1023       | Alice Johnson  | Laptop     | 2        | 899.99   | 
| 1045       | Bob Smith      | Smartphone | 1        | 599.99   | 
| 1078       | Carol Lee      | Tablet     | 3        | 299.99   | 
| 1101       | David Wong     | Headphones | 4        | 199.99   | 

```{r}
CustomerID = c(1023, 1045, 1078, 1101)
CustomerName = c("Alice Johson", "Bob Smith", "Carol Lee", "David Wong")
Product = c("Laptop", "Smartphone", "Table", "Headphones")
Quantity = c(2,1,3,4)
UnitPrice = c(899.99, 599.99, 299.99, 199.99)
Sales_Data = data.frame(CustomerID,CustomerName, Product, Quantity, UnitPrice)
```

### Q5: Display the structure of the data frame
- Use `str()` to display the structure of `sales_data`.
```{r}
str(Sales_Data)
```

### Q6: Convert data types in the data frame
- Ensure the following column types:
- CustomerID integer
- CustomerName character
- Product factor
- Quantity integer
- UnitPrice numeric
```{r}
class(CustomerID)
class(CustomerName)
class(Product)
class(Quantity)
class(UnitPrice)
class(TotalPrice)
CustomerID = as.integer(CustomerID)
Product = as.factor(Product)
Quantity = as.integer(Quantity)
```

### Q7: Compute the average purchase amount
- create a new column called TotalPrice to sales_data multiplying the columns. code: sales_data$TotalPrice = sales_data$Quantity * sales_data$UnitPrice
- After in the same chunk, calculate the average `TotalPrice` for all transactions from sales_data.
```{r}
Sales_Data$Total_Price = Sales_Data$Quantity * Sales_Data$UnitPrice
mean(Sales_Data$Total_Price)
```

### Q8: Find the most frequently purchased product
- Display the count of each product in `sales_data`.
```{r}
table(Sales_Data$Product)
```

### Q9: Display the number of purchases per customer
```{r}
table(Sales_Data$Product, Sales_Data$CustomerID)
```

### Q10: Create a subset for the sales_data where total purchase is greater than $500
- Identify and display transactions where `TotalPrice` is greater than 500 and use subset to create the new dataframe
called sales_500
```{r}
sales_500 <- subset(Sales_Data, Total_Price > 500)
```

### Q11: Display transactions where Quantity is greater than 2
- Display transactions where `Quantity` is more than 2. Hint: Use rows, columns and throw the sales_data$Quantity > 2 in
the rows section of sales_data[rows,columns]
```{r}
Sales_Data[Sales_Data$Quantity > 2,]
```

### Q12: Display the first row of `sales_data` for all columns. Hint: Rows, Columns
```{r}
Sales_Data[1,]
```

### Q13: Extract the UnitPrice of the third row-- Retrieve and display the `UnitPrice` for the third row.
```{r}
Sales_Data[3,"UnitPrice"]
```

### Q14: Update the Quantity of the second row to 5
- Modify the `Quantity` value for the second row to 5.
- Display the updated data frame. (just type sales_data)
```{r}
Sales_Data[2,"Quantity"] <- 5
Sales_Data$Total_Price = Sales_Data$Quantity * Sales_Data$UnitPrice
Sales_Data
```

### Q15: Replace the customer name of the fourth row with "Doe"
- Change the customer name in the fourth row to " Doe ".
- Display the updated data frame.
```{r}
Sales_Data[4,"CustomerName"] <- "Doe"
Sales_Data
```



**ADD THE QUESTIONS BELOW AND A CHUNK FOLLOWING EACH TO THE LAB2.RMD**





